---
title: "Assignment 6"
<<<<<<< HEAD
author: "Charles Lang, Edited by Hsiao Yang" 
=======
author: "Charles Lang" Edited by Hsiao Yang
>>>>>>> 78a66c218b82b823759c202e5ed3dcd0a4bfb079
date: "11/16/2016"
output: html_document
---
#Addignment 6

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

##Part I

#Packages
```{r}
library(rpart)
<<<<<<< HEAD
# install.packages("
library(tidyverse)
=======
install.packages("rpart.plot")
>>>>>>> 78a66c218b82b823759c202e5ed3dcd0a4bfb079
library(rpart.plot)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("MOOC1.csv", header = TRUE)

M2 <- read.csv("MOOC2.csv", header = TRUE)

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame. Which variables should you use?

c.tree1 <- rpart(as.factor(certified) ~ grade + assignment, method = "class", data = M1)

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)

#Plot your tree

rpart.plot(c.tree1)
post(c.tree1, file = "tree1.ps", title = "MOOC") 

#This creates a pdf image of the tree

```
##Comment
Variables that we will include is grade and assignment so we can prune the tree. If we include forum posts, there is only two nodes so then we will not be able to prune the tree for part II of this assignment.

##Part II

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Complexity Parameter" and represents the cost to error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
c.tree2 <- prune(c.tree1, cp = 0.058182 )#Set cp to the level at which you want the tree to end

#Visualize this tree and compare it to the one you generated earlier

printcp(c.tree2)
rpart.plot(c.tree2)

post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree
```

#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

```{r}
M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict2 <- predict(c.tree2, M2, type = "class")

T1 <- table(M2$certified, M2$predict1)

T2 <- table(M2$certified, M2$predict2)
<<<<<<< HEAD

#Tree 1 True Results
T1TR <- (T1[2,2]+T1[1,1])/sum(T1)

=======

#Tree 1 True Results
T1TR <- (T1[2,2]+T1[1,1])/sum(T1)

>>>>>>> 78a66c218b82b823759c202e5ed3dcd0a4bfb079
#Tree 2 True Results
T2TR <- (T2[2,2]+T2[1,1])/sum(T2)

```
#Comment
The second tree has a lower error rate. This is because the True Negative and True Positive for tree 2 is 53.63% while tree 1 is 21.86% so that means the second tree has a higher accuracy vesus the first tree.

##Part III

Choose a data file from the (University of Michigan Open Data Set)[https://github.com/bkoester/PLA/tree/master/data]. Choose an outcome variable that you would like to predict. Build two models that predict that outcome from the other variables. The first model should use raw variables, the second should feature select or feature extract variables from the data. Which model is better according to the cross validation metrics?

```{r}

<<<<<<< HEAD
R.pre <- read.csv("student.record.csv", header = TRUE)
R <- R.pre %>% 
  filter(HSGPA <= 4)

c.tree3 <- rpart(HSGPA ~ ., data = R)

printcp(c.tree3)

R$predict1 <- predict(c.tree3, R)

T1 <- R %>% 
  select(HSGPA, predict1) %>% 
  mutate(
    residual = HSGPA - predict1,
    squared.residual = residual^2
  )

sqrt(mean(T1$squared.residual, na.rm = T))

c.tree3$variable.importance
R.select <- R %>% 
  select(HSGPA, ADMIT_TERM, MAJOR1_DESCR, MAJOR1_DEPT)

c.tree4 <- rpart(HSGPA ~ ., data = R.select)

R.select$predict1 <- predict(c.tree4, R.select)

T2 <- R.select %>% 
  select(HSGPA, predict1) %>% 
  mutate(
    residual = HSGPA - predict1,
    squared.residual = residual^2
  )

sqrt(mean(T2$squared.residual, na.rm = T))


=======
R <- read.csv("student.record.csv", header = TRUE)

c.tree3 <- rpart(as.factor(HSGPA) ~., method = "class", data = R)

printcp(c.tree3)

>>>>>>> 78a66c218b82b823759c202e5ed3dcd0a4bfb079
```
##Comment
First model prediction = 1.322
Second model prediction = 1.296

The second model that selected Admit_Term, Major1_Descr, and Major1_Dept has a better prediction since the prediction is lower than the first model, which is closer to 0.


### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.