---
title: "Assignment 6"
author: "Yiwen Ma"
date: "11/16/2016"
output: html_document
---
#Addignment 6

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

##Part I

#Packages
```{r}
library(rpart)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("MOOC1.csv", header = TRUE)

M2 <- read.csv("MOOC2.csv", header = TRUE)

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame. Which variables should you use?

c.tree <- rpart(as.factor(certified) ~ forum.posts+grade+assignment, method = "class", data = M1)
c.tree1 <- rpart(as.factor(certified) ~ grade+assignment, method = "class", data = M1)

#Check the results from the classifcation tree using the printcp() command
printcp(c.tree)
printcp(c.tree1)


#Plot your tree
post(c.tree, file = "tree.ps", title = "MOOC") 
post(c.tree1, file = "tree1.ps", title = "MOOC") #This creates a pdf image of the tree

```

When including 3 variables(c.tree), the desicion tree is overwhelmed by forum.posts, who has a larger range from 0-200. Thus, choosing only grade and assignment as variables in the desicion tree(c.tree1) is much better and get a more balanced tree.

##Part II

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Complexity Parameter" and represents the cost to error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
c.tree2 <- prune(c.tree1, cp = 0.058182 )#Set cp to the level at which you want the tree to end
printcp(c.tree2)
#Visualize this tree and compare it to the one you generated earlier

post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree
```

#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

```{r}
M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict2 <- predict(c.tree2, M2, type = "class")

table1 <- table(M2$certified, M2$predict1)

table2 <- table(M2$certified, M2$predict2)

accuracy1 <- sum(diag(table1))/sum(table1)
accuracy2 <- sum(diag(table2))/sum(table2)


accuracy1
accuracy2
```

Tthe first prediction model has an accuracy of 0.2186, the second prediction model has an accuracy of 0.5363, which is much better. Thus, the second tree has a lower error rate.

##Part III

Choose a data file from the (University of Michigan Open Data Set)[https://github.com/bkoester/PLA/tree/master/data]. Choose an outcome variable that you would like to predict. Build two models that predict that outcome from the other variables. The first model should use raw variables, the second should feature select or feature extract variables from the data. Which model is better according to the cross validation metrics?

```{r}
library('randomForest')

data2 <- read.csv("student.record.csv", header = TRUE)
#Choose numerical value from data 2
data3 <- data2[,c(3:9,13)]

# Deal with NA
data4 <- na.omit(data3)
data5 <- subset(data4, data4$MAJOR1_DESCR != "")
```


# model 1 Use raw variables

```{r}
model1 <- rpart(as.factor(MAJOR1_DESCR) ~HSGPA+ LAST_ACT_ENGL_SCORE+LAST_ACT_MATH_SCORE+LAST_ACT_READ_SCORE+LAST_ACT_SCIRE_SCORE+LAST_ACT_COMP_SCORE+as.factor(SEX), method = "class", data = data5)

printcp(model1)

# CP nsplit rel error  xerror      xstd
# 1 0.02168      0   1.00000 1.00000 0.0012575
# 2 0.01000      1   0.97832 0.97832 0.0014301
```

# model 2 Use trasformed data

```{r}
library(stringr)
#Conclude the MAJOR1_DESCR into larger group
data5$DIP<-word(data5$MAJOR1_DESCR, -1)
data5$DIP[which(data5$DIP=="Fresh")]<-"BS"
data5$DIP[which(data5$DIP=="ChemE")]<-"BSE"
data5$DIP[which(data5$DIP=="BSChem")]<-"BS"
data5$DIP[which(data5$DIP=="Chem")]<-"BS"
data5$DIP[which(data5$DIP=="CompS")]<-"BSE"
data5$DIP[which(data5$DIP=="ElecE")]<-"BSE"
data5$DIP[which(data5$DIP=="CS")]<-"BSE"
data5$DIP[which(data5$DIP=="Aero")]<-"BSE"
data5$DIP[which(data5$DIP=="Sc")]<-"BSE"
data5$DIP[which(data5$DIP=="E")]<-"BSE"

#Select the top 5 category for prediction
s1 <- sort(table(data5$DIP),decreasing = TRUE)[1:5]
data6 <- data5[which(data5$DIP %in% rownames(s1)),]

#sacle the dataset
library(dplyr)
library(rpart.plot)
data6<-filter(data6,HSGPA != 0) 
ind <- sapply(data6, is.numeric)
data7<- as.data.frame(lapply(data6[ind], scale))
data7 <- cbind(data7, data6[,8:9])

#model2
model2 <- rpart(as.factor(DIP) ~HSGPA+ LAST_ACT_ENGL_SCORE+LAST_ACT_MATH_SCORE+LAST_ACT_READ_SCORE+LAST_ACT_SCIRE_SCORE+LAST_ACT_COMP_SCORE+as.factor(SEX), method = "class", data = data7)

printcp(model2)
#xerror:0.89035

rpart.plot(model2)
```

# model 3 Use random forest

```{r}

model3 <- randomForest(as.factor(DIP) ~HSGPA+ LAST_ACT_ENGL_SCORE+LAST_ACT_MATH_SCORE+LAST_ACT_READ_SCORE+LAST_ACT_SCIRE_SCORE+LAST_ACT_COMP_SCORE+SEX, data = data7)

#number of trees
plot(model3)

print(model3)
#OOB estimate of  error rate: 52.67%

```

# Conclusion

According to 3 models above,

Model1 : used raw data as input and got nearly 97.832% of error rate.

Model2 : After combined different types of major into larger group(eg:BA,BE,BSE), and scaled the ACT score, the error rate decrese to 90%, which is much lower than the model1.

This model is used to predict the diploma which student choosed based on the ACT score. The result seems reasonable. Student with lower ACT math score will choose Bachelor of Arts. Students with both high math score and high highschool GPA will tend to choose Bachelor of Sciences. Besides, gender also plays an important role in major, male will prefer BSE (Engineering), and female with high math score will choose BA and BS.

Model3 : After using random forest to further improve the model, the error rate droped to 52.67%.

### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.