---
title: "Assignment 6"
author: "Charles Lang"
date: "11/16/2016"
output: html_document
---
#Assignment 6-Euki Sawada

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

#Packages
```{r}
library(rpart)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("MOOC1.csv", header = TRUE)

M2 <- read.csv("MOOC2.csv", header = TRUE)

```

#Decision tree
```{r}
# Q1: Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame. Which variables should you use?

# A1: I used grade and assignment because forum.posts was too overpowering, resulting in a tree with one node instead of two. 

c.tree1 <- rpart(certified ~ grade + assignment, method="class", data=M1, control = rpart.control(minsplit = 1))

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)

#Plot your tree

post(c.tree1, file = "tree1.ps", title = "MOOC") #This creates a pdf image of the tree

#Explore tree with forum.posts after scaling (not used later)
c.tree1scaled <- rpart(certified ~ scale(forum.posts) + scale(grade) + scale(assignment), method="class", data=M1, control = rpart.control(minsplit = 1))

printcp(c.tree1scaled)

post(c.tree1scaled, file = "tree1scaled.ps", title = "MOOC Scaled")
```

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Cost Complexity" and represents the cost in error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
c.tree2 <- prune(c.tree1, cp = 0.058182)#Set cp to the level at which you want the tree to end (pruned to one node)

#Visualize this tree and compare it to the one you generated earlier

post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree
```

# Q2: Now use both the original tree and the pruned tree to make predictions about the students in the second data set. Which tree has a lower error rate?

# A2:  The pruned tree had a higher accuracy (53.63%) than the original tree (21.86%) which supports the idea that the latter was overfitted and the former is more generalizable.

```{r}
M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict2 <- predict(c.tree2, M2, type = "class")

table(M2$certified, M2$predict1)

table(M2$certified, M2$predict2)

sum(M2$certified == M2$predict1)/length(M2$certified)*100

sum(M2$certified == M2$predict2)/length(M2$certified)*100

```