---
title: "Assignment 6"
author: "Qiyang(Minnie)Lin"
date: "11/16/2019"
output: html_document
---
#Addignment 6

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

##Part I

#Packages
```{r}
library(rpart)
library(rpart.plot)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("MOOC1.csv", header = TRUE)

M2 <- read.csv("MOOC2.csv", header = TRUE)

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame. Which variables should you use?

M1$forum.posts_Scale<- scale(M1$forum.posts)
M1$grade_Scale<- scale(M1$grade)
M1$assignment_Scale<-scale(M1$assignment)

#M1$ScaleAll<- scale(M1$forum.posts,M1$grade, M1$assignment)

c.tree1 <- rpart(as.factor(certified) ~ grade + assignment, method="class", data=M1)

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)

#Plot your tree

post(c.tree1, file = "tree1.ps", title = "MOOC") #This creates a pdf image of the tree
rpart.plot(c.tree1)

```

##Part II

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Complexity Parameter" and represents the cost to error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
c.tree2 <- prune(c.tree1, cp = 0.058182 )#Set cp to the level at which you want the tree to end (see the CP table)

#Visualize this tree and compare it to the one you generated earlier

post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree
rpart.plot(c.tree2)
```

#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

```{r}
M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict2 <- predict(c.tree2, M2, type = "class")

table(M2$certified, M2$predict1)

table(M2$certified, M2$predict2)

#tree2 has lower error rate. because the sum of "no"+"no" and "yes"+"yes" are 5363 > 2186 in tree1. 

#error rate in predict1
M2$count1<- ifelse(M2$predict1 == M2$certified, 1, 0)
M2$count2<- ifelse(M2$predict2 == M2$certified, 1, 0)

mean(M2$count1) #0.2186
mean(M2$count2) #0.5363

#tree2 has higher prediction accuracy. 

```


##Part III

Choose a data file from the (University of Michigan Open Data Set)[https://github.com/bkoester/PLA/tree/master/data]. Choose an outcome variable that you would like to predict. Build two models that predict that outcome from the other variables. The first model should use raw variables, the second should feature select or feature extract variables from the data. Which model is better according to the cross validation metrics?

```{r}

D1<- read.csv("student.course.csv")
#because the original dataset is too big, so I randomly select 1000 of them. 
D2<- D1[sample(nrow(D1), 1000), ]

#Model 1 that include all the raw variables to predict GPAO. 
c.tree3 <- rpart(GPAO ~ ANONID + CATALOG_NBR + SUBJECT + GRD_PTS_PER_UNIT + DIV + ANON_INSTR_ID + TERM, data = D2)

printcp(c.tree3)

rpart.plot(c.tree3)

#Model 2: feature selection/extractrion
D3<-dplyr::select(D2, GPAO, GRD_PTS_PER_UNIT, TERM, CATALOG_NBR, ANONID, ANON_INSTR_ID)
D3<- sapply(D3, as.numeric )
D4<- data.frame(scale(D3))

#use pca 
pca <- prcomp(D4, scale = TRUE)

pca1<- data.frame(pca$x)

D5<- dplyr::bind_cols(D2, pca1)

#c.tree4<- rpart(GPAO ~ GRD_PTS_PER_UNIT + TERM , data= D2)
c.tree5<- rpart(GPAO ~ PC1 + PC2, data = D5)
printcp(c.tree5)

rpart.plot(c.tree5)


#Discussion: Both two models are used to predict GPAO. The first model use all raw variables in the data, creating a 8-split decision tree that has cross validation error of 0.728 and standard error of 0.127. The second model use PC1 and PC2 to predit GPAO, creating a 8-split decision tree that has cross validation error of 0.417 and standard error of 0.067. According to the resluts, the second model is better which contains less cross validaiton error. 


```


### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.