---
title: "Assignment 6"
author: "Charles Lang"
date: "11/16/2016"
output: html_document
programmer: "Li Li"
last modified: "Nov 20 2018"
---
#Addignment 6

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

#Packages
```{r}
library(rpart)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("MOOC1.csv", header = TRUE)

M2 <- read.csv("MOOC2.csv", header = TRUE)

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame. Which variables should you use?

##Answer: I would assume that those who get higher grades in final exam and assignments would have higher probabilities of getting certified. So I would choose "grade + assignment" 2 variables in this case.
##Note: previously, when choosing all three variables, only forum.posts are unsed by R in tree construction.

c.tree1 <- rpart(certified ~ grade + assignment, method="class", data=M1)

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)

#Plot your tree

post(c.tree1, file = "tree1.ps", title = "MOOC") #This creates a pdf image of the tree

```

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Cost Complexity" and represents the cost in error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
c.tree2 <- prune(c.tree1, cp=0.2)#Set cp to the level at which you want the tree to end
#Question: is there any rule for setting the cp to a certain level? I chose to put "0.2" here just to get rid of the third layer.

#Visualize this tree and compare it to the one you generated earlier

post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree

#Answer: this tree has less nodes than the previous tree, and the error rate in predicting the "small sample" is higher in this tree than in previous tree. This increase in error rate reflects that our current model do not "over-fit" the  data, as the previous model may have done. 
```

#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

```{r}
M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict2 <- predict(c.tree2, M2, type = "class")

table(M2$certified, M2$predict1)

table(M2$certified, M2$predict2)

#Answer: 
##for tree1: the error rate=(24+7790)/10000=0.7814
##for tree2(pruned): the error rate=(1184+3453)/10000=0.4637
##therefore, the pruned tree2 does a better job in predicting on the "wild sample".
```
