---
title: "Assignment 6"
author: "Fernando Carnauba"
date: "11/25/2019"
output: html_document
---
#Assignment 6

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

##Part I

#Packages
```{r}
library(rpart)
library(party)
library(rattle)
library(RColorBrewer)
library(dplyr)
library(readr)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("~/R/HUDK/assignment6/MOOC1.csv", header = TRUE)
M2 <- read.csv("~/R/HUDK/assignment6/MOOC2.csv", header = TRUE)

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame. Which variables should you use?

c.tree1 <- rpart(certified ~ forum.posts + grade +assignment, method="class", data=M1)

#Check the results from the classifcation tree using the printcp() command
printcp(c.tree1)

#Plot your tree
# post(c.tree1, file = "tree1.ps", title = "MOOC") #This creates a pdf image of the tree
fancyRpartPlot(c.tree1, caption = NULL)

```
# ANSWER: The algorithm automatically selects only "forum.posts" as a relevant variable for the three. It is actually quite striking to see in the tree plot that the threshold of 80 forum posts divides the sample almost perfectly. NEITHER of he 27% of students with less than 80 posts got a certificate, while vitually ALL (99%) of the students who posted 80 or more times did get a certificate. It seems quite reasonable, then, to only use forum.posts for building the tree. 


##Part II

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Complexity Parameter" and represents the cost to error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
c.tree2 <- prune(c.tree1, cp = 0)#Set cp to the level at which you want the tree to end

#Visualize this tree and compare it to the one you generated earlier
#post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree
fancyRpartPlot(c.tree2, caption = NULL)
```

#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

ANSWER: Nothing different between the trees, since the first one already had only one variable taken into account, with an almost perfect match. I tried really hard to go back and set parameters to force the fisrt three to actually use the three variables, but I just couldn't make it work. So, for the sake of my learning experience with comparing error rates, I will create a third three, that uses GRADES and ASSIGNMENT to predict certification (three #3) and I will compare it with tree #1 that uses only the variable FORUM.POSTS.

```{r}
# my new tree (created for "pedagogical purposes"):
c.tree3 <- rpart(certified ~  grade +assignment, method="class", data=M1)

M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict3 <- predict(c.tree3, M2, type = "class")

table(M2$certified, M2$predict1)

table(M2$certified, M2$predict3)

fancyRpartPlot(c.tree3, caption = NULL)

```

Analysis:
First tree (uses only "forum.posts"):
- Predicts about 45% of the "true NO" correctly.
- Preducts about 55% if the "true YES" correctly.

Third tree (uses only "grade" and "assignment"):
- Predicts about 99% of the "true NO" correctly.
- Preducts about 1% if the "true YES" correctly.

The new tree seems to be very biased towards predicting students not to have a certificate. Probably due to a very "high" grading threshold of 7.5. 

##Part III

Choose a data file from the (University of Michigan Open Data Set)[https://github.com/bkoester/PLA/tree/master/data]. Choose an outcome variable that you would like to predict. Build two models that predict that outcome from the other variables. The first model should use raw variables, the second should feature select or feature extract variables from the data. Which model is better according to the cross validation metrics?

ANSWER: 
I am actually very interested in gender biases in education. So I decided to try to predict SEX as an outcome variable from the other information in the dataset. For example, do higher grades predict gender? Does the type of major predict gender? Does doing two majors predict gender? 

Model A - High school GPA, SAT scores, ACT scores
Model B - Same data as above, but now calculated as AVERAGES BY MAJOR, and also a dummy variable for engineering course. By using this, I want to see if men/women bundle in more/less selective courses, in term of previous SAT/ACT scores.[I have decided ex-post to eliminate HSGPA, because the tree was using only this data into the analysis. I don't know exactly way, but using the dummy for engineering was leading to an error in the rpart routine, and the tree coudn't be calculated. So I gave up on that variable as well]


```{r}
# Read data
UM <- as.data.frame(read_csv("https://raw.githubusercontent.com/fcarnauba/PLA/master/data/student.record.csv"))
UM <- select(UM,SEX,HSGPA, 
             LAST_SATI_VERB_SCORE,
             LAST_SATI_MATH_SCORE,
             LAST_SATI_TOTAL_SCORE,
             LAST_ACT_ENGL_SCORE,
             LAST_ACT_MATH_SCORE,
             LAST_ACT_READ_SCORE,
             LAST_ACT_SCIRE_SCORE,
             LAST_ACT_COMP_SCORE,
             MAJOR1_DESCR)


#create averages of GPA, SAT and ACT for each major:
UM <- UM %>%
      group_by(MAJOR1_DESCR) %>%
      mutate(mean.HSGPA = mean(HSGPA),
         mean.SATI_VERB_SCORE  = mean(LAST_SATI_VERB_SCORE),
         mean.SATI_MATH_SCORE  = mean(LAST_SATI_MATH_SCORE),
         mean.SATI_TOTAL_SCORE = mean(LAST_SATI_TOTAL_SCORE),
         mean.ACT_ENGL_SCORE   = mean(LAST_ACT_ENGL_SCORE),
         mean.ACT_MATH_SCORE   = mean(LAST_ACT_MATH_SCORE),
         mean.ACT_READ_SCORE   = mean(LAST_ACT_READ_SCORE),
         mean.ACT_SCIRE_SCORE  = mean(LAST_ACT_SCIRE_SCORE),
         mean.ACT_COMP_SCORE   = mean(LAST_ACT_COMP_SCORE))

# create engineering variable (ended up not using, because it bugged the rpart routine)
UM$eng <- as.factor(ifelse(grepl("eng", UM$MAJOR1_DESCR), "engineer", "other"))

# FIRST MODEL
c.treeA <- rpart(SEX ~ HSGPA  + 
                 + LAST_SATI_VERB_SCORE
                 + LAST_SATI_MATH_SCORE
                 + LAST_SATI_TOTAL_SCORE
                 + LAST_ACT_ENGL_SCORE
                 + LAST_ACT_MATH_SCORE
                 + LAST_ACT_READ_SCORE
                 + LAST_ACT_SCIRE_SCORE
                 + LAST_ACT_COMP_SCORE,
                 method="class", data=UM)

# Plot First model
fancyRpartPlot(c.treeA, caption = NULL)


# SECOND MODEL
c.treeB <- rpart(SEX ~
                  mean.SATI_MATH_SCORE
                 + mean.SATI_TOTAL_SCORE
                 + mean.ACT_ENGL_SCORE
                 + mean.ACT_MATH_SCORE
                 + mean.ACT_READ_SCORE
                 + mean.ACT_SCIRE_SCORE
                 + mean.ACT_COMP_SCORE,
                 method="class", data=UM)

# Plot Second model
fancyRpartPlot(c.treeB, caption = NULL)



```

## MODEL1.
One very interesting feature is that the model used both SAT and ACT mathematics scores, even though there were many other scores available. A mathematics bias towards men is a well-researched phenomenon in the US, and we might me seing some of it here. While a high GPE is an indicator of a higher probability of being a female, math scores both in the ACT and SAT go in the opposite direction.


## MODEL2.
We can now see a "math bias" in terms of the average math (and science) SAT and ACT scores in the majors. Overall, majors in which those grades are higher have less women.

Now let's look into the errors of each mdoel

```{r}
# my new tree (created for "pedagogical purposes"):
UM$predictA <- predict(c.treeA, UM, type = "class")
UM$predictB <- predict(c.treeB, UM, type = "class")

table(UM$SEX, UM$predictA)

table(UM$SEX, UM$predictB)


```

Model A provides a much better prediction, with 66% of women correctly pinned by the tree, and 44% of men correctly predicted. Model B basically assigns "everyone" as a man, with nearly 100% of women wrongly predicted, and 100% of men correctly predicted.   


### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.