---
title: "Assignment 6"
author: "Charles Lang"
date: "11/16/2016"
output: html_document
---
#Addignment 6

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

#Packages
```{r}
library(rpart)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.table("MOOC1.csv", sep = ",", header = TRUE)

M2 <- read.table("MOOC2.csv", sep = ",", header = TRUE)

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame.

c.tree1 <- rpart(certified ~ grade + forum.posts, method="class", data=M1, control=rpart.control(minsplit = 1, minbucket = 1, cp=0.001))

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)

#         CP nsplit rel error xerror     xstd
#1 0.0040850      0   1.00000 1.0000 0.062466
#2 0.0036765      7   0.97059 1.0588 0.063790
#3 0.0029412     11   0.95588 1.1029 0.064731
#4 0.0024510     36   0.88235 1.2059 0.066761
#5 0.0019608     44   0.86275 1.2108 0.066852
#6 0.0014006     49   0.85294 1.2255 0.067123
#7 0.0012255     56   0.84314 1.2451 0.067477
#8 0.0010000     72   0.82353 1.2549 0.067651

#Plot your tree

post(c.tree1, file = "treemooc1.ps", title = "Certified: 1 - Yes, 2 - No, 3 - ?")

#terrible looking tree. but I think that was what was expected. 

```

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Cost Complexity" and represents the cost in error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
#prune back to 4th
c.tree2 <- prune(c.tree1, cp = 0.0024510)

#Visualize this tree and compare it to the one you generated earlier

post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree
```

#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

```{r}
M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict2 <- predict(c.tree2, M2, type = "class")

table(M2$certified, M2$predict1)
#       no  yes
#  no   196 1884
#  yes  798 7122
table(M2$certified, M2$predict2)
#       no  yes
#  no   145 1935
#  yes  539 7381

# wrong predictions over total number for each prediction model
(1935+539)/(145+1935+539+7381)
# predict2 0.2474

(1884+798)/(196+1884+798+7122)
# predict1 0.2682

#Prediction2 has lower error percentage, so better in this case with this data. 
#Deep breath. 
```
