---
title: "Assignment 6"
author: "Charles Lang"
date: "11/16/2016"
output: html_document
---
#Assignment 6

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

#Packages
```{r}
#setwd("C:/Users/Magdalena Bennett/Dropbox/PhD Columbia/Fall 2016/Core Methods in EDM/assignment6b")

library(rpart)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("MOOC1.csv")

M2 <- read.csv("MOOC2.csv")

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame.

M1$certified <- as.numeric(M1$certified) - 1

c.tree1 <- rpart(certified ~ ., method = "class", data = M1,
                 control = rpart.control(minsplit = 1, minbucket = 1, cp = 0.001))

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)

#Plot your tree

post(c.tree1, file = "ctree1.ps", title = "Certification prediction")

par(xpd = TRUE)
plot(c.tree1, compress = TRUE)
text(c.tree1, use.n = TRUE)

```

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Cost Complexity" and represents the cost in error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}

#I decided to cut the tree at the third level.

c.tree2 <- prune(c.tree1, cp = 0.0036765)

#Visualize this tree and compare it to the one you generated earlier

post(c.tree2, file = "ctree2.ps", title = "MOOC") #This creates a pdf image of the tree

par(xpd = TRUE)
plot(c.tree2, compress = TRUE)
text(c.tree2, use.n = TRUE)
```

#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

```{r}
M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict2 <- predict(c.tree2, M2, type = "class")

table(M2$certified, M2$predict1)

table(M2$certified, M2$predict2)

```

- We can analyze the proportion of correct predictions in each tree:

```{r}
correct1 <- sum(diag(table(M2$certified, M2$predict1)))/sum(table(M2$certified, M2$predict1))

correct2 <- sum(diag(table(M2$certified, M2$predict2)))/sum(table(M2$certified, M2$predict2))

correct1

correct2
```

- The second model works better as prediction, probably because it's not overfitting the data as much as the first one.

- However, sometimes false negatives might be more costly than false positives, so we might want to compare this rate instead of an overall rate:

```{r}

falseneg1 <- table(M2$certified, M2$predict1)[2,1]/sum(table(M2$certified, M2$predict1)[,1])

falseneg2 <- table(M2$certified, M2$predict2)[2,1]/sum(table(M2$certified, M2$predict2)[,1])

falseneg1

falseneg2

```

- In this case, the second model has a slightly higher rate of false negatives (out of all the negative outcomes it predicts, 80% of them are incorrectly predicted), but the difference is small enough, compared to the overall prediction error improvement with respect to the first model, so we might still choose the first one.