---
title: "Assignment 6"
author: "Charles Lang"
date: "11/16/2016"
output: html_document
---
#Addignment 6

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

##Part I

#Packages
```{r}
library(rpart)
library(tidyr)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("MOOC1.csv", header = TRUE)

M2 <-read.csv("MOOC2.csv",header=TRUE)

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame. Which variables should you use?
#using grade for course exam and course assignment.if we include forum.post it will dominant the prediction. Do not know why.
#M1%>% mutate_at(c(2,3,4), funs(c(scale(.))))
# or M1[c(2,3,4)]<-scale(M1[c(2,3,4)])
c.tree1 <- rpart(factor(certified)~grade+assignment,method="class", data=M1)

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)

#Plot your tree

post(c.tree1, file = "tree1.ps", title = "MOOC") #This creates a pdf image of the tree

```

##Part II

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Complexity Parameter" and represents the cost to error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
c.tree2 <- prune.rpart(c.tree1, cp = 0.058182 )#Set cp to the level at which you want the tree to end
#cp2=0.058182
#Visualize this tree and compare it to the one you generated earlier

post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree
#grade variable is not longer exist. 
```

#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

```{r}


M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict2 <- predict(c.tree2, M2, type = "class")

table(M2$certified, M2$predict1)

table(M2$certified, M2$predict2)
# The second tree has a lower error rate. However, based on the cross validation metrics,the original 2 split should be better, however, it has lower accruacy. Should try different model.
#install.packages("neuralnet")
library(neuralnet)
nn=neuralnet(certified~forum.posts+grade+assignment,data=M1, hidden=3,act.fct = "logistic",
                linear.output = FALSE)
plot(nn)
Predict=compute(nn,M2)
predict3<-Predict$net.result
predict3<-predict3[,1]
predict3 <- ifelse(predict3>0.5, "yes", "no")
M2<-cbind(M2,predict3)
table(M2$certified,M2$predict3)
#seems this one have similar accuracy with predict2 (2 nods one). still not very accurate.
#install.packages("ranger")
#install.packages("randomForest")
#install.packages("pROC")
library(pROC)
library(ROCR)
library(randomForest)
library(ranger)
rf<-randomForest(formula =certified~grade+assignment,data=M1)
M2$predict4<-predict(rf,M2)
table(M2$certified, M2$predict4)
#accuracy is even lower......
rgr1 <- ranger(factor(certified)~.,data = M1,
               num.trees = 1000,respect.unordered.factors = T)
prediction <- predict(rgr1,M2,type="response")
#pred<-prediction(prob,M2$certified)#numeric variable

M2$predict5<- prediction$predictions
table(M2$certified,M2$predict5)

#rgr2 <- ranger(factor(certified)~.,data = M1,
               #num.trees = 1000,respect.unordered.factors = T,probability = T)
#mean(ifelse(M2$certified==M2$predict5,1,0))
#auc<-auc(roc(M2$certified,M2$predict5))
#auc<- performance(prediction,measure='auc')@y.values[[1]]

```

##Part III

Choose a data file from the (University of Michigan Open Data Set)[https://github.com/bkoester/PLA/tree/master/data]. Choose an outcome variable that you would like to predict. Build two models that predict that outcome from the other variables. The first model should use raw variables, the second should feature select or feature extract variables from the data. Which model is better according to the cross validation metrics?

```{r}
M3 <- read.csv("student.course.csv", header = TRUE)
M4<-read.csv("student.record.csv")

library(dplyr)
library(rpart)
# I decided to anlyze student.course dataset.
# run a correlation plot to figure out relationship

library(corrplot)

#M3_sample<-M3[sample(nrow(M3), 10000), ]
#M3_sample<-select(M3_sample,-SUBJECT,-DIV)
#M3_sample<-scale(M3_sample)
#M3_sample<-as.data.frame(M3_sample)
#run prediction model with all variable except ID
c.tree3 <- rpart(GPAO~GRD_PTS_PER_UNIT+TERM+CATALOG_NBR, method="anova", data=M3)
printcp(c.tree3)
post(c.tree3, file = "tree3.ps", title = "GPAO")
#run correlation plot to figure out most correlated variables, I decided to choose the most related variables, GRD_PTS_PER_UNIT and TERM.
COR<-select(M3,-SUBJECT,-DIV)
COR<-scale(COR)
COR<-cor(COR)
corrplot<-corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
c.tree4 <- rpart(GPAO~GRD_PTS_PER_UNIT+TERM,method="anova", data=M3)
printcp(c.tree4)
post(c.tree4, file = "tree4.ps", title = "GPAO")

#predict something else, maybe subject
c.tree5<-rpart(factor(DIV)~GPAO+GRD_PTS_PER_UNIT,method="class",data=M3)
c.tree6<-rpart(factor(SUBJECT)~GPAO+GRD_PTS_PER_UNIT,method="class",data=M3)
printcp(c.tree5) 
printcp(c.tree6)
#Seems like no correlation between DIV and SUBUJECT with GPAO and GRD_PTS_PER_UNIT??
#By comparing cross validation metrices, the second model (with all related variables) has both lowest xerror & xstd
#so I prefer using the original model with 4 splits. (However, slightly different)
#If xerror and standard deviation is not that different everysplit, should I choose the smallest one? or should balance the number of split(less is better) with xerror and xstd as well.
#might be using PCA (pc component to make a prediction model?)

#seems like student course data not a good data to do prediction in different model. The only determint variable is grade_points_per_unit.

M4<-select(M4,SEX,MAJOR1_DESCR,HSGPA,LAST_ACT_ENGL_SCORE,LAST_ACT_MATH_SCORE,LAST_ACT_READ_SCORE,LAST_ACT_SCIRE_SCORE,LAST_ACT_COMP_SCORE)
M4<-na.omit(M4)
c.tree7<-rpart(SEX~LAST_ACT_ENGL_SCORE+LAST_ACT_MATH_SCORE+LAST_ACT_READ_SCORE+LAST_ACT_SCIRE_SCORE+LAST_ACT_COMP_SCORE,method="class",data=M4)
post(c.tree7, file = "tree7.ps", title = "SEX")
printcp(c.tree7)
c.tree8<-rpart(SEX~HSGPA+LAST_ACT_ENGL_SCORE+LAST_ACT_READ_SCORE+LAST_ACT_SCIRE_SCORE+LAST_ACT_COMP_SCORE,method="class",data=M4)
printcp(c.tree8)
#comparing by two models, the first model(c.tree7)has both less split,xerror and xstd. So I prefer the first model.
```


### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.