---
title: "Assignment 6"
author: "Charles Lang"
date: "11/16/2016"
output: html_document
---
#Addignment 6

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

#Packages
```{r}
library(rpart)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("~/Desktop/FALL_2016/Assignment 6/MOOC1.csv")

M2 <- read.csv("~/Desktop/FALL_2016/Assignment 6/MOOC2.csv")

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame.

c.tree1 <- rpart(certified ~ forum.posts + grade + assignment,method="class", data = M1, control=rpart.control(minsplit=1,minbucket=1,cp=0.001))

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)
          CP nsplit rel error xerror     xstd
1  0.0044563      0   1.00000 1.0000 0.062466
2  0.0039216     20   0.90196 1.1422 0.065531
3  0.0036765     36   0.83333 1.1961 0.066577
4  0.0032680     40   0.81863 1.2794 0.068079
5  0.0029412     53   0.77451 1.3578 0.069371
6  0.0028011     65   0.73529 1.3578 0.069371
7  0.0024510    100   0.61765 1.4216 0.070339
8  0.0016340    198   0.37255 1.4608 0.070900
9  0.0012255    250   0.27451 1.5686 0.072310
10 0.0010000    262   0.25980 1.5735 0.072370

#Plot your tree

post(c.tree1, file = "tree.ps", title = "certified yes/no")

```

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Cost Complexity" and represents the cost in error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
c.tree2 <- prune(c.tree1, cp =  0.0016340)

in here you should use your own cp
          CP nsplit rel error xerror     xstd
1  0.0044563      0   1.00000 1.0000 0.062466
2  0.0039216     20   0.90196 1.1422 0.065531
3  0.0036765     36   0.83333 1.1961 0.066577
4  0.0032680     40   0.81863 1.2794 0.068079
5  0.0029412     53   0.77451 1.3578 0.069371
6  0.0028011     65   0.73529 1.3578 0.069371
7  0.0024510    100   0.61765 1.4216 0.070339
8  0.0016340    198   0.37255 1.4608 0.070900
9  0.0012255    250   0.27451 1.5686 0.072310
10 0.0010000    262   0.25980 1.5735 0.072370
#cp3
c.tree3 <- prune(c.tree1, cp = 0.0036765)
printcp(c.tree3)
post(c.tree3, file = "tree3.ps", title = "MOOC")
#cp2
c.tree4 <- prune(c.tree1, cp = 0.0039216)
printcp(c.tree4)
post(c.tree4, file = "tree4.ps", title = "MOOC")
#Visualize this tree and compare it to the one you generated earlier
post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree
```

#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

```{r}
M2$predict1 <- predict(c.tree1, M2, type = "class")
M2$predict2 <- predict(c.tree4, M2, type = "class")

table(M2$certified, M2$predict1)
table(M2$certified, M2$predict2)

        no  yes
  no   432 1648
  yes 1680 6240
  (1680+1648)/(432+1648+1680+6240)
  0.3328
add up all four and divde with yesno and noyes add up (error)
     
        no  yes
  no    64 2016
  yes  239 7681
  (2016+239)/(64+2016+239+7681)
  0.2255
First one was original cp and second one has the lower error rate which used cp2. Here adding less node means lesser error rate and maybe adding the more nodes means more over fitting.
```
