---
title: "Assignment 6_Group 6"
author: "Mahshad D"
date: "12/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Task I
* You will be working in groups on a quantified student project
* Download the *Sensor Kinetics Pro* app to your [iOS](https://itunes.apple.com/us/app/sensor-kinetics-pro/id623633248?mt=8) or [Android](https://play.google.com/store/apps/details?id=com.innoventions.sensorkineticspro&hl=en) device (or similar gyroscope measuring app)
* Split into subgroups with each subgroup investigating one measures available in the app
* Each subgroup shopuld report back aafter 20min with what they have learned  

##Please see https://docs.google.com/document/d/1bkCk6XsSL_e085Si-h1YvBbpkEhOxEwpKKUIGNYZ7VA/edit


### Task II
* In consultation with your group select a physical activity that: a) can be measured using the app, b) has a concrrete, sontinuous, countable outcome and c) some members of the group are experts and some are novices at - If you are looking for inspiration you can find previous student projects below.
* Collect measurements for all members of the group performing the task using the app
* Create a Github repo where you can store and share your data
* Using K-means, can you identify the novices from the experts using only the app data?
* Visualize your results using ggplot2

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(readr)

D1 <- read_csv("Group 6  data.csv")
D2 <- read_csv("distance.csv")
D3 <- merge(D1, D2)
D3 <- D3[,-1]
D3 <- scale(D3)
fit <- kmeans(D3, 2)
D4 <- data.frame(D3, fit$cluster)
ggplot(D4, aes(distance, time, colour = fit.cluster)) + geom_point() + xlab("Distance") + ylab("Time")

```
#Yes, the experts are the data points with the lowest distance (highest stability) and longest time; all in all, the four data points in the middle of the graph are the experts.


### Task III
* Devise five questions on a five point Likert scale that ask members of your group about their experience with your chosen task
* Collect data from all members of your group
* Run a PCA analysis on the data
* What does PC1 represent in your analysis?
* Visualize the PC1 results for each member of your group

```{r}

```

### Task IV
* If you could only choose one thing to predict the score for each member of your group (raw data variable, a cluster or a PC) what would it be and why?
* Create a visualization that supports your conclusion
* Based on your conclusion devise an intervention that would help novice members of your group imporve their performance
```{r}


```

