---
title: "Lizarov,Anna.Assignment#6"
author: "Anna Lizarov"
date: "November 20, 2018"
output: html_document
---

#Packages
```{r}
library(rpart)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("MOOC1.csv", header = TRUE)

M2 <- read.csv("MOOC2.csv", header = TRUE)

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame. Which variables should you use?

#I used variables grade and assignment.  

c.tree1 <- rpart(certified~grade+assignment, method="class", data=M1)

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)

#Plot your tree

post(c.tree1, file = "tree1.ps", title = "MOOC") #This creates a pdf image of the tree
```

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Cost Complexity" and represents the cost in error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
c.tree2 <- prune(c.tree1, cp = 0.058182)#Set cp to the level at which you want the tree to end

#Visualize this tree and compare it to the one you generated earlier

post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree
```

```{r}
# Comparison of the trees

#Answer: The pruned tree only predicts the certified based on the students' average grade for the course assignments while the original tree predicts certfied based on both the students' average grade for the course assignments and the students' average grade for the course exam. The error rate based on the students' average grade for the course assignments is the same for both trees. 
```


#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

```{r}
M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict2 <- predict(c.tree2, M2, type = "class")

table(M2$certified, M2$predict1)

table(M2$certified, M2$predict2)
```

```{r}
#Calculation of an error rate of the original tree
#There are 10000 students

(24+7790)/10000  
```

```{r}
#Calculation of an error rate of the pruned tree
(1184+3453)/10000
```

```{r}
# Answer: The pruned tree seems to have a lower error rate of 46.37% whereas the original tree has a high error rate of 78.14%. As one can see, there is a substantial difference between the error rates of the pruned tree and the original tree.  
```

