---
title: "H5_data_cleaning"
author: "Paolo Rivas"
date: "12/11/2020"
output: html_document
---

## Data Cleaning H5
Here we have two mayor problems. The first one, there are different amounts of data points per users. Second, our time variable is shown as a character with an excesive ammount of milliseconds. None of those are useful for a uniform measurement.

Stept 1: import Dataframe an summarize information

```{r}

library(tidyverse)

DF <- read_csv("~/Documents/raw_data_H5_all.csv")
summary(DF)

```
Step 2: transform time(s) into timeframe data.
```{r}
class(DF$`Time (s)`) 
```
```{r}
#this format works for formating time correctly into seconds and milliseconds
t <- 0.00250;
t2<- format(as.POSIXct(t, origin = "2020-01-12", tz = "America/chicago"), "%H:%M:%OS3")
class(t2)
#Now lets format this new character into time
install.packages("chron")
library(chron)
x <- chron(times=t2)
class(x)

#### Now lets do it for our data

t3 <- format(as.POSIXct(DF$`Time (s)`, origin = "2020-01-12", tz = "America/New_York"), "%H:%M:%S")
DF$time_formated <- chron(times=t3)
class(DF$time_formated) #We have a time daframe!
```
## Now, lets create time average dplyr with a double group_by

```{r}
library(dplyr)
DF_clean <- DF %>%
  group_by(Participant, time_formated) %>%
  summarise(avg_x = mean(`Acceleration x (m/s^2)`),avg_y = mean(`Acceleration y (m/s^2)`), avg_z = mean(`Acceleration z (m/s^2)`) )

```
## Exporting the data to CSv
```{r}
write.csv(df, "Clean_data_H5.csv")
```

